### 2016-01-19 Manuscript/SW_early_life_telomeres.Rmd
Post_evolution major changes

#Set some options
options(na.action='na.fail')
opts_knit$set(root.dir = '..')
opts_chunk$set(echo=FALSE,
               warning=FALSE,
               message=FALSE,
               fig.path='../Figures/',
               fig.width= 4,
               fig.height= 8,
               fig.show='asis',
               dev=c('png','pdf'))

@@

Lewis Spurgin: lewisspurgin@gmail.com; Kat Bebbington: K.Bebbington@uea.ac.uk Eleanor Fairfield: E.Fairfield@uea.ac.uk; Jan Komdeur: j.komdeur@rug.nl; Terry Burke: t.a.burke@sheffield.ac.uk; Hannah Dugdale: h.l.dugdale@rug.nl; David Richardson: david.richardson@uea.ac.uk

@@

There is evidence from a range of taxa that the greatest rate of telomere loss occurs in early life [e.g. @Frenck1998; @Haussmann2003], and that the extent of this telomere shortening is influenced by the conditions experienced during that period [@Price2013; @Monaghan2014; @Nettle2015; @Reichert2015]. Importantly, early-life telomere dynamics have been associated with both short-term and late-life survival [@Heidinger2012; @Boonekamp2014], and with other parameters such as cognition [@Nettle2015a]. However, to our knowledge no studies have simultaneously analysed how telomere dynamics, early-life conditions and late-life survival are all related in a natural setting. Moreover, how early-life telomere dynamics vary over spatial and temporal scales is not known.

The longitudinal study (since 1986) of the Seychelles warbler (*Acrocephalus sechellensis*) population on Cousin Island provides an excellent model system for studying senescence in the wild [reviewed in @Hammers2015]. Due to the isolated nature of the study population and intensive field monitoring, we have unusually comprehensive survival data and tissue samples spanning many years (see Methods, below). Ecological conditions and warbler population density on Cousin vary across space and time due to weather-induced changes in foliage cover and food availability [@VandeCrommenacker2011a]. Seychelles warblers remain on their natal territories for at least six months, and variation in the oxidative stress experienced by individuals is associated with natal  territory quality [@VandeCrommenacker2011a]. However, neither early-life nor adult survival appear to be associated with natal territory quality or natal local density [@Brouwer2006; @Hammers2013]. Facultative cooperative breeding occurs in the Seychelles warbler [@Komdeur1994a; @Richardson2003a], and the presence of helping subordinates (but not non-helping subordinates) in the natal territory is associated with increased survival later in life [@Brouwer2012]. Lastly, we have an established protocol for assessing absolute telomere length in this species [@Barrett2012], and telomere length predicts survival independently of age in adult Seychelles warblers, suggesting that telomeres act as a biomarker of cost [@Barrett2013a]. Thus, we have an excellent system in which to assess the costs of different social and environmental conditions experienced early in life, and to assess the later-life consequences of early-life conditions.

In this study we test the prediction that telomeres link early-life environmental variation to late-life fitness. Because telomere dynamics are expected to reflect individual-level variation in the costs of early-life experiences, they may allow a more sensitive analysis of the effects of early-life environmental variation than would be possible with a direct comparison of how survival is affected by the early-life environment. With this in mind, we predict that: i) individuals raised in higher quality conditions will have longer telomeres and reduced telomere shortening, and ii) longer telomeres and lower rates of telomere shortening in early life are associated with greater survival.

@@

#####Early-life telomere length and age
Mean ($\pm$ standard error) telomere length in first-year birds was `r round2(mean(FlSAall$TLKB))` $\pm$ `r round2(se(FlSAall$TLKB))` kb. This was lower than the mean for one-year-old birds based on cross-sectional data (`r round2(mean(subset(dd,Age==1)$TLKB))` $\pm$ `r round2(se(subset(dd,Age==1)$TLKB))` kb). However, when we considered age classes within the first year of life separately, we found substantial age-related variation in telomere length (F = `r round2(anova(lm(TLKB~Fledged,data = juv))[1,4])`; P = `r round3(anova(lm(TLKB~Fledged,data = juv))[1,5])`). Nestlings had longer telomeres than any other age group, but there was an apparent increase in average telomere length from the fledgling to adult stages (Fig. 1A). However, a longitudinal analysis of telomere loss showed that nestlings and fledglings both lost telomeric DNA early in life, and that there was no difference in the rate of telomere loss between the two age classes (F = `r round2(anova(lm(TROC~Fledged,data = Loss))[1,4])`; P = `r round2(anova(lm(TROC~Fledged,data = Loss))[1,5])`; Fig. 1B), suggesting that the cross-sectional results represent selective disappearance.

#####Among-cohort variation in early-life telomere dynamics
Early-life telomere length varied significantly among breeding seasons (one-way ANOVA, F = `r round2(anova(lm(TLKB~factor(FieldPeriodID),data = juv))[1,4])`; P = `r round3(anova(lm(TLKB~factor(FieldPeriodID),data = juv))[1,5])`), and this variation in mean telomere length over breeding seasons was positively related to island-wide food availability (linear regression, R^2^ = `r round2(summary(lm(Insect~TLKBmean,data=juvseason,na.action=na.exclude))$r.squared)`; F = `r round2(anova(lm(Insect~TLKBmean,data = juvseason,na.action=na.exclude))[1,4])`; P = `r round3(anova(lm(Insect~TLKBmean,data = juvseason,na.action=na.exclude))[1,5])`; Fig. 2A). Running the regression of cohort-level telomere length and food availability separately for the two age classes revealed a positive, significant relationship in fledglings (R^2^ = `r round2(summary(lm(Insect~TLKBmean,data=flseason,na.action=na.exclude))$r.squared)`; F = `r round2(anova(lm(Insect~TLKBmean,data=flseason,na.action=na.exclude))[1,4])`; P = `r round3(anova(lm(Insect~TLKBmean,data=flseason,na.action=na.exclude))[1,5])`), but no relationship in nestlings (R^2^ = `r round2(summary(lm(Insect~TLKBmean,data=chickseason,na.action=na.exclude))$r.squared)`; F = `r round2(anova(lm(Insect~TLKBmean,data=chickseason,na.action=na.exclude))[1,4])`; P = `r round3(anova(lm(Insect~TLKBmean,data=chickseason,na.action=na.exclude))[1,5])`).

@@

#####Within-cohort variation in early-life telomere dynamics
The top model explaining within-season variation in early-life telomere length contained age class, tarsus length and the number of helpers present in the natal territory, as well as interactions between age class and both tarsus length and the number of helpers (Table S1). This model was much better supported than the null model ($\Delta$AICc = `r round2(AICc(lm(cenTL~1,data=juvall))-AICc(lm(cenTL~Fledged*(cenTarsus+cenHelper),data=juvall)))`). Both tarsus length and the number of helpers had high relative importance in the top model set (Fig. 3A), and the model containing these variables explained a reasonable amount of variation in telomere length (R^2^ = `r round2(summary(lm(cenTL~Ageclass*(cenTarsus+cenHelper),data=juvall))$r.squared)`). Telomere length was positively related to the number of helpers in nestlings, but not in fledglings (Fig. 3B). Tarsus length was negatively related to telomere length in nestlings, but not in fledglings (Fig. 3C).

For the subset of individuals for which we had longitudinal data, we tested whether the factors previously identified as being related to early-life telomere length (i.e. tarsus length and number of helpers) were also associated with differences in rates of early-life telomere loss. Due to a limited sample size, we considered all age classes together for this analysis. We found that none of these variables were associated with differences in telomere loss (all P > 0.05).

#####Early-life telomere dynamics and survival
At the population-level, there was no relationship between lifespan and telomere length (R^2^ = `r round3(summary(lm(Lifespan~TLKBmean,data=juvseason,na.action=na.exclude))$r.squared)`; F = `r round2(anova(lm(Lifespan~TLKBmean,data=juvseason,na.action=na.exclude))[1,4])`; P = `r round2(anova(lm(Lifespan~TLKBmean,data=juvseason,na.action=na.exclude))[1,5])`; Fig. 4A). Testing this relationship separately for each age class revealed that there was no age-specific relationship between lifespan and cohort-level telomere length (both P > 0.2). There was a negative relationship between population-level telomere loss and lifespan, although this was marginally non-significant (R^2^ = `r round3(summary(lm(Lifespan~TROCmean,data=loss.season,na.action=na.exclude))$r.squared)`; F = `r round2(anova(lm(Lifespan~TROCmean,data=loss.season,na.action=na.exclude))[1,4])`; P = `r round2(anova(lm(Lifespan~TROCmean,data=loss.season,na.action=na.exclude))[1,5])`; Fig. 4B).

For the individual-based survival analysis, a log-normal survival model best fitted the Seychelles warbler data (AIC = `r round2(flexsurvreg(with(FlSA,Surv(Lifespan+1,Died))~ 1, dist="lnorm")$AIC)`, AIC of next best distribution (Weibull) = `r round2(flexsurvreg(with(FlSA,Surv(Lifespan+1,Died))~ 1, dist="weibull")$AIC)`). A survival model including all first-year birds showed that early-life telomere length had a positive effect on survival, but this relationship was marginally non-significant (estimate = `r round3(flexsurvreg(with(juvall,Surv(Lifespan+1,Died)~cenTL), dist="lnorm")$res[3,1], lessthan = F)`, CI = `r round3(flexsurvreg(with(juvall,Surv(Lifespan+1,Died)~cenTL), dist="lnorm")$res[3,2], lessthan = F)`-`r round3(flexsurvreg(with(juvall,Surv(Lifespan+1,Died)~cenTL), dist="lnorm")$res[3,3], lessthan = F)`). Running this analysis between age classes revealed differential survival effects. Nestling telomere length had no effect on survival  (estimate = `r round3(flexsurvreg(with(chickall,Surv(Lifespan+1,Died)~cenTL), dist="lnorm")$res[3,1], lessthan = F)`, CI = `r round3(flexsurvreg(with(chickall,Surv(Lifespan+1,Died)~cenTL), dist="lnorm")$res[3,2], lessthan = F)`-`r round3(flexsurvreg(with(chickall,Surv(Lifespan+1,Died)~cenTL), dist="lnorm")$res[3,3], lessthan = F)`; Fig. 5A), but telomere length at the fledgling stage was significantly related to survival (estimate = `r round3(flexsurvreg(with(FlSAall,Surv(Lifespan+1,Died)~cenTL), dist="lnorm")$res[3,1], lessthan = F)`, CI = `r round3(flexsurvreg(with(FlSAall,Surv(Lifespan+1,Died)~cenTL), dist="lnorm")$res[3,2], lessthan = F)`-`r round3(flexsurvreg(with(FlSAall,Surv(Lifespan+1,Died)~cenTL), dist="lnorm")$res[3,3], lessthan = F)`). This effect was positive, with increased survival later in life associated with longer telomeres (Fig. 6B). Finally, using the longitudinal data, we found no effect of early-life telomere shortening on subsequent survival, although again the estimate was positive, and confidence intervals only marginally overlapped zero (estimate = `r round2(flexsurvreg(with(Loss,Surv(Lifespan+1,Died)~cenTROC), dist="lnorm")$res[3,1])`, CI = `r round3(flexsurvreg(with(Loss,Surv(Lifespan+1,Died)~cenTROC), dist="lnorm")$res[3,2], lessthan = F)`-`r round3(flexsurvreg(with(Loss,Surv(Lifespan+1,Died)~cenTROC), dist="lnorm")$res[3,3], lessthan = F)`).

>>>>

#Set some options
options(na.action='na.fail')
opts_knit$set(root.dir = '..')
opts_chunk$set(echo=FALSE,
               warning=FALSE,
               message=FALSE)

@@

**Correspondence:** Lewis Spurgin: lewisspurgin@gmail.com; David Richardson: david.richardson@uea.ac.uk

@@

There is evidence from a range of taxa that the greatest rate of telomere loss occurs in early life [e.g. @Frenck1998; @Haussmann2003], and that the extent of this telomere shortening is influenced by the conditions experienced during that period [@Price2013; @Monaghan2014; @Nettle2015; @Reichert2015]. Importantly, early-life telomere dynamics have been associated with both short-term and late-life survival [@Heidinger2012; @Boonekamp2014], and with other parameters such as cognition [@Nettle2015a]. However, few studies have simultaneously analysed how telomere dynamics, early-life conditions and late-life survival are all related in a natural setting. Moreover, how early-life telomere dynamics vary over spatial and temporal scales is poorly understood.

The longitudinal study (since 1986) of the Seychelles warbler (*Acrocephalus sechellensis*) population on Cousin Island provides an excellent system for studying senescence in the wild [reviewed in @Hammers2015]. Due to the isolated nature of the study population and intensive field monitoring, we have unusually comprehensive survival data and tissue samples spanning many years (see Methods, below). Ecological conditions and warbler population density on Cousin vary across space and time due to weather-induced changes in foliage cover and food availability [@VandeCrommenacker2011a]. Seychelles warblers remain on their natal territories for at least six months, and variation in the oxidative stress experienced by individuals is associated with natal  territory quality [@VandeCrommenacker2011a]. However, neither early-life nor adult survival appear to be associated with natal territory quality or natal local density [@Brouwer2006; @Hammers2013]. Facultative cooperative breeding occurs in the Seychelles warbler [@Komdeur1994a; @Richardson2003a], and the presence of helping subordinates (but not non-helping subordinates) in the natal territory is associated with increased survival later in life [@Brouwer2012]. Lastly, we have an established protocol for assessing absolute telomere length in this species [@Barrett2012], and telomere length predicts survival independently of age in adult Seychelles warblers, suggesting that telomeres act as a biomarker of cost [@Barrett2013a]. Thus, we have an excellent system in which to assess the costs of different social and environmental conditions experienced early in life, and to assess the later-life consequences of early-life conditions.

In this study we examine how telomeres link early-life environmental variation to late-life fitness in the Seychelles warbler. Because telomere dynamics are expected to reflect individual-level variation in the costs of early-life experiences, they may allow a more sensitive analysis of the effects of early-life environmental variation than would be possible with a direct comparison of how survival is affected by the early-life environment. With this in mind, we first conduct an exploratory analysis of how the environmental and social factors experienced in early life affect telomere dynamics. We then test the hypothesis that longer telomeres and lower rates of telomere shortening in early life are associated with greater survival.

@@

#####Early-life telomere length and age
Mean ($\pm$ standard error) telomere length in first-year birds was longer than in adults of any age (Fig. 1A) We also found substantial age-related variation in telomere length within the first year of life, with longer telomeres in nestlings compared to fledglings (ANOVA; F = `r round2(anova(lm(TLKB~Fledged,data = juv))[1,4])`; P = `r round3(anova(lm(TLKB~Fledged,data = juv))[1,5])`; Fig. 1B).

#####Among-cohort variation in early-life telomere dynamics
Early-life telomere length varied significantly among breeding seasons (one-way ANOVA, F = `r round2(anova(lm(TLKB~factor(FieldPeriodID),data = juv))[1,4])`; P = `r round3(anova(lm(TLKB~factor(FieldPeriodID),data = juv))[1,5])`), with a tendency toward shorter telomeres in more recent years (linear regression, R^2^ = `r round2(summary(lm(TLKBmean~Tarsus,data=juvseason,na.action=na.exclude))$r.squared)`; F = `r round2(anova(lm(Insect~TLKBmean,data = juvseason,na.action=na.exclude))[1,4])`; P = `r round3(anova(lm(Insect~TLKBmean,data = juvseason,na.action=na.exclude))[1,5])`; Fig. 2). Variation in mean telomere length over breeding seasons was positively related to island-wide food availability (linear regression, R^2^ = `r round2(summary(lm(TLKBmean~Tarsus,data=juvseason,na.action=na.exclude))$r.squared)`; F = `r round2(anova(lm(Insect~TLKBmean,data = juvseason,na.action=na.exclude))[1,4])`; P = `r round3(anova(lm(Insect~TLKBmean,data = juvseason,na.action=na.exclude))[1,5])`; Fig. 2A). Running the regression of cohort-level telomere length and food availability separately for the two age classes revealed a positive, significant relationship in fledglings (R^2^ = `r round2(summary(lm(Insect~TLKBmean,data=flseason,na.action=na.exclude))$r.squared)`; F = `r round2(anova(lm(Insect~TLKBmean,data=flseason,na.action=na.exclude))[1,4])`; P = `r round3(anova(lm(Insect~TLKBmean,data=flseason,na.action=na.exclude))[1,5])`), but no relationship in nestlings (R^2^ = `r round2(summary(lm(Insect~TLKBmean,data=chickseason,na.action=na.exclude))$r.squared)`; F = `r round2(anova(lm(Insect~TLKBmean,data=chickseason,na.action=na.exclude))[1,4])`; P = `r round3(anova(lm(Insect~TLKBmean,data=chickseason,na.action=na.exclude))[1,5])`).

@@

#####Within-cohort variation in early-life telomere dynamics
The top model explaining within-season variation in early-life telomere length contained age class, tarsus length and the number of helpers present in the natal territory, as well as interactions between age class and both tarsus length and the number of helpers (Table S1). This model was much better supported than the null model ($\Delta$AICc = `r round2(AICc(lm(cenTL~1,data=juvall))-AICc(lm(cenTL~Fledged*(cenTarsus+cenHelper),data=juvall)))`; R^2^ = `r round2(summary(lm(cenTL~Ageclass*(cenTarsus+cenHelper),data=juvall))$r.squared)`). Both tarsus length and the number of helpers had high relative importance in the top model set (Fig. 3A). Telomere length was positively related to the number of helpers in nestlings, but not in fledglings (Fig. 3B). Tarsus length was negatively related to telomere length in nestlings, but not in fledglings (Fig. 3C).

For the subset of individuals for which we had longitudinal data, we tested whether the factors previously identified as being related to early-life telomere length (i.e. tarsus length and number of helpers) were also associated with differences in rates of early-life telomere loss. Due to a limited sample size, we considered all age classes together for this analysis. The number of helpers in the natal territory was negatively associated to the rate of telomere loss (). We found no relationship between tarsus length and telomere loss ().

#####Early-life telomere dynamics and survival

library(survival)
m1 <- coxph(with(juv,Surv(Lifespan+1,Died)~LogTL+GroupSize+Fledged+frailty(LayYear)))

A log-normal survival model best fitted the Seychelles warbler data (AIC = `r round2(flexsurvreg(with(juv,Surv(Lifespan+1,Died))~ 1, dist="lnorm")$AIC)`, AIC of next best distribution (Weibull) = `r round2(flexsurvreg(with(juv,Surv(Lifespan+1,Died))~ 1, dist="weibull")$AIC)`). A survival model including all first-year birds showed that early-life telomere length had a positive effect on survival, but this relationship was marginally non-significant (estimate = `r round3(flexsurvreg(with(juvall,Surv(Lifespan+1,Died)~TLKB+LayYear), dist="lnorm")$res[3,1], lessthan = F)`, CI = `r round3(flexsurvreg(with(juvall,Surv(Lifespan+1,Died)~TLKB), dist="lnorm")$res[3,2], lessthan = F)`-`r round3(flexsurvreg(with(juvall,Surv(Lifespan+1,Died)~TLKB), dist="lnorm")$res[3,3], lessthan = F)`). Running this analysis between age classes revealed differential survival effects. Nestling telomere length had no effect on survival  (estimate = `r round3(flexsurvreg(with(chickall,Surv(Lifespan+1,Died)~TLKB), dist="lnorm")$res[3,1], lessthan = F)`, CI = `r round3(flexsurvreg(with(chickall,Surv(Lifespan+1,Died)~TLKB), dist="lnorm")$res[3,2], lessthan = F)`-`r round3(flexsurvreg(with(chickall,Surv(Lifespan+1,Died)~TLKB), dist="lnorm")$res[3,3], lessthan = F)`; Fig. 5A), but telomere length at the fledgling stage was significantly related to survival (estimate = `r round3(flexsurvreg(with(FlSAall,Surv(Lifespan+1,Died)~TLKB), dist="lnorm")$res[3,1], lessthan = F)`, CI = `r round3(flexsurvreg(with(FlSAall,Surv(Lifespan+1,Died)~TLKB), dist="lnorm")$res[3,2], lessthan = F)`-`r round3(flexsurvreg(with(FlSAall,Surv(Lifespan+1,Died)~TLKB), dist="lnorm")$res[3,3], lessthan = F)`). This effect was positive, with increased survival later in life associated with longer telomeres (Fig. 6B). Finally, using the longitudinal data, we found no effect of early-life telomere shortening on subsequent survival (estimate = `r round2(flexsurvreg(with(Loss,Surv(Lifespan+1,Died)~cenTROC), dist="lnorm")$res[3,1])`, CI = `r round3(flexsurvreg(with(Loss,Surv(Lifespan+1,Died)~cenTROC), dist="lnorm")$res[3,2], lessthan = F)`-`r round3(flexsurvreg(with(Loss,Surv(Lifespan+1,Died)~cenTROC), dist="lnorm")$res[3,3], lessthan = F)`).

### 2016-01-19 Manuscript/SW_early_life_telomeres_tables_and_figures.Rmd
Post_evolution major changes

```{r Figure 1}
######################
#1A
######################

ddFig1A <- ddply(dd,
                 .(Fledged),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = se(TLKB))

Fig1A <- ggplot(ddFig1A,
       aes(x = Fledged,
           y = TLKBmean)) +
  ylab('Telomere length (kb)') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.1,0.9),
        legend.title = element_blank(),
        legend.text = element_text(size = 15)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  annotate('text',label = 'A',x=0.6,y=6.2)

######################
#1B
######################

ddFig1B <- ddply(Loss,
                 .(Fledged),
                 summarize,
                 TROCmean = mean(TROC),
                 TROCse = se(TROC))

Fig1B <- ggplot(ddFig1B,
       aes(x = Fledged,
           y = TROCmean)) +
  ylab('Corrected telomere loss') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.1,0.9),
        legend.title = element_blank(),
        legend.text = element_text(size = 15)) +
  geom_pointrange(aes(ymin=TROCmean-TROCse,ymax=TROCmean+TROCse)) +
  annotate('text',label = 'B',x=0.6,y=0.0018)

pdf('Figures/Figure 1.pdf',
    width = 4,
    height = 8,
    paper = 'a4')
multiplot(Fig1A,Fig1B)
dev.off()

@@

####################

Fig2A <- ggplot(juvseason,
       aes(x = Insect,
           y = TLKBmean,
           col = Season)) +
  ylab('Telomere length (kb)') + 
  xlab('Food availability') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  stat_smooth(method = 'lm',se = T,col = 'black') +
  scale_color_manual(values = c('black',grey(0.5))) +
  annotate('text',label = 'A',x=-1.5,y=8)

@@

temp <- loss.season[complete.cases(loss.season),]

####################################
Fig2B <- ggplot(temp,
       aes(x = Insect,
           y = TROCmean,
           col = Season)) +
  geom_pointrange(aes(min=TROCmean-TROCse,ymax=TROCmean+TROCse)) +
  ylab('Corrected telomere loss') + 
  xlab('Food availability') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  stat_smooth(method = 'lm',se = T,  col = 'black') +
  scale_color_manual(values = c('black',grey(0.5))) +
  annotate('text',label = 'B',x=-1.5,y=0.0075)

pdf('Figures/Figure 2.pdf',
    width = 4,
    height = 8,
    paper = 'a4')
multiplot(Fig2A,Fig2B)
dev.off()

@@

######################
#3A
######################

#Run the model
lm2 <- lm(cenTL ~ Fledged*(cenHelper + cenNonHelper + Sex + cenTarsus  + cenTQ),
            data = juvall)

@@

lm3 <- dredge(lm2)
lmavg <- model.avg(lm3,subset = delta <=6)
myimp <- importance(lmavg) #Relative importance of each explanatory variable

@@

Fig3A <- forestplot(table2) +
  ylim(c(-1,1)) +
  annotate('text',label = 'A',x=0.6,y=1)

######################
#3B
######################

temp <- juv
temp$HelperF <- ifelse(juv$Helper>0,'Helpers present','Helpers absent')

ddFig3B <- ddply(temp,
                 .(HelperF,Fledged),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = se(TLKB))

Fig3B <- ggplot(ddFig3B,
       aes(x = HelperF,
           y = TLKBmean,
           col = Fledged,
           lty = Fledged)) +
  ylab('Telomere length (kb)') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none',
        legend.title = element_blank(),
        legend.text = element_text(size = 15)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse),
                  position = position_dodge(width = 0.2)) +
  scale_color_manual(values = c('black',grey(0.5))) +
  annotate('text',label = 'B',x=0.5,y=8)

######################
#3C
######################

Fig3C <- ggplot(juv,
                aes(x = Tarsus,
                    y = TLKB,
                    col = Fledged,
                    lty = Fledged)) +
  ylab('Telomere length (kb)') + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_point(size=3) +
  stat_smooth(method = 'lm',se = T) +
  scale_color_manual(values = c('black',grey(0.5))) +
  annotate('text',label = 'C',x=19,y=15)

pdf('Figures/Figure 3.pdf',
    width = 8,
    height = 8,
    paper = 'a4')
multiplot(Fig3A,Fig3B,Fig3C,layout=matrix(c(1,1,2,3),2,2,byrow=TRUE))
dev.off()

@@

```{r Figure 4,fig.height=8,fig.width=4}

Fig4A <- ggplot(juvseason,
                aes(x = TLKBmean,
                    y = Lifespan,
                    col = Season)) +
  geom_pointrange(aes(ymin=Lifespan-Lifespanse,ymax=Lifespan+Lifespanse)) +
  geom_errorbarh(aes(xmin = TLKBmean-TLKBse,xmax = TLKBmean+TLKBse)) +
  theme_classic() +
  ylab('Remaining life') + 
  xlab('Telomere length (kb)') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none') +
  scale_color_manual(values = c('black',grey(0.5))) +
  stat_smooth(method = 'lm',se = T,col = 'black',fullrange = T) +
  annotate('text',x=2.5,y=10,label='A')

Fig4B <- ggplot(loss.season,
                aes(x = TROCmean,
                    y = Lifespan,
                    col = Season)) +
  geom_pointrange(aes(ymin=Lifespan-Lifespanse,ymax=Lifespan+Lifespanse)) +
  geom_errorbarh(aes(xmin = TROCmean-TROCse,xmax = TROCmean+TROCse)) +
  theme_classic() +
  ylab('Remaining life') + 
  xlab('Corrected telomere loss') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 11), 
        axis.text.y = element_text(size = 11), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none') +
  scale_color_manual(values = c('black',grey(0.5))) +
  stat_smooth(method = 'lm',se = T,col = 'black',fullrange = T) +
  annotate('text',x=-0.004,y=12.5,label='B')

pdf('Figures/Figure 4.pdf',
    width = 4,
    height = 8,
    paper = 'a4')
multiplot(Fig4A,Fig4B)
dev.off()

@@

```{r Figure 5,fig.height=8,fig.width=4}

s1 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = chicks)
Fig5A <- ggsurv(s1,plot.cens=F,lty.est=c(1,2)) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  xlim(c(0,15)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  scale_colour_manual(name = 'TL',
                      values = c('black',grey(0.5))) +
  guides(colour=F) +
  annotate('text',x=0.5,y=1.1,label='A')

###########################################
s2 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = FlSAall)
Fig5B <- ggsurv(s2,plot.cens=F,lty.est=c(1,2)) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  xlim(c(0,15)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  scale_colour_manual(name = 'TL',
                      values = c('black',grey(0.5))) +
  guides(colour=F) +
  annotate('text',x=0.5,y=1.1,label='B')

pdf('Figures/Figure 5.pdf',
    width = 4,
    height = 8,
    paper = 'a4')
multiplot(Fig5A,Fig5B)
dev.off()

>>>>

```{r Figure 1}

####################
temp <- subset(juvseason,Season == 'Summer')
temp <- subset(temp,FieldPeriodID != 27)

Fig1 <- ggplot(temp,
       aes(x = CatchYear,
           y = TLKBmean,
           col = Season)) +
  ylab('Log telomere length (kb)') + 
  xlab('Lay Year') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  geom_line()+
  scale_color_manual(values = c('black',grey(0.5)))

pdf('Figures/Figure 1.pdf',
    width = 6,
    height = 4,
    paper = 'a4')
Fig1
dev.off()

@@

####################
Fig1 <- ggplot(dd,
       aes(x = Fledged,
           y = LogTL,
           col = Season)) +
  ylab('Telomere length (kb)') + 
  xlab('Lay Year') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  geom_line()+
  scale_color_manual(values = c('black',grey(0.5)))

@@

pdf('Figures/Figure 1.pdf',
    width = 6,
    height = 4,
    paper = 'a4')
Fig1
dev.off()

@@

#####################

#Run the model
lm2 <- lmer(LogTL ~ Fledged + Tarsus + Helper + GroupSize + Sex +Tarsus + BodyMass  + cenTQ  + (1|LayYear) + (1|PlateID),
            data = juv,
            REML = FALSE)

@@

lm3 <- dredge(lm2)
lm3 <- subset(lm3,delta <=6)
lmavg <- model.avg(lm3,fit=T)
myimp <- importance(lmavg) #Relative importance of each explanatory variable

@@

Fig3 <- forestplot(table2) +
  annotate('text',label = 'A',x=0.6,y=0.1)

pdf('Figures/Figure 2.pdf',
    width = 8,
    height = 8,
    paper = 'a4')
Fig3
dev.off()

@@

```{r Figure 3,fig.height=8,fig.width=4}

s1 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = chickall)
Fig3A <- ggsurv(s1,plot.cens=F,lty.est=c(1,2)) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  xlim(c(0,15)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  scale_colour_manual(name = 'TL',
                      values = c('black',grey(0.5))) +
  guides(colour=F) +
  annotate('text',x=0.5,y=1.1,label='A')

###########################################
s2 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = FlSAall)
Fig3B <- ggsurv(s2,plot.cens=F,lty.est=c(1,2)) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  xlim(c(0,15)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  scale_colour_manual(name = 'TL',
                      values = c('black',grey(0.5))) +
  guides(colour=F) +
  annotate('text',x=0.5,y=1.1,label='B')

pdf('Figures/Figure 3.pdf',
    width = 4,
    height = 8,
    paper = 'a4')
multiplot(Fig3A,Fig3B)
dev.off()

@@
